Characterising representable positive relation algebras with Priestley duality

### 2023-04-02 duality/Text.tex
proof sketch with a couple of todos

\author{Daniel Rogozin}
\date{}
\title{Characterising representable positive relation algebras via Priestley duality}

@@

@@

\begin{enumerate} Let $\mathcal{L}$ be a distributive lattice and let $\mathcal{X}$ be a Priestley space:
\item $\mathcal{L} \hookrightarrow \mathcal{L}^{+} = (2^{\operatorname{Spec}(\mathcal{L})}, \cap, \cup, \emptyset, \operatorname{Spec}(\mathcal{L}))$,
\item $\mathcal{L} \cong \operatorname{ClOp}(\operatorname{Spec}(\mathcal{L}))$,
\item $\mathcal{X} \cong \operatorname{Spec}(\operatorname{ClOp}(\mathcal{X}))$,
\item The categories of Priestley spaces and bounded distributive lattices are dually equivalent.
\end{enumerate}
\end{fact}

@@

\section{Spectral spaces for positive relation algebras}

@@

\begin{lemma} Let $\mathcal{R}$ be a positive relation algebra, then

\begin{enumerate}
\item
\item 
\end{enumerate}
\end{lemma}

\section{Complete representability}

\begin{enumerate}
\item $\mathcal{L}$ is completely representable iff $\mathcal{L}$ has a distinguishing set of complete, completely prime filters,
\item $(\mathcal{L}_+)^{+}$ is completely representable.
\end{enumerate}

\end{theorem}

\subsection{Completely representable positive relation algebras}

@@

\begin{theorem}
Let $\mathcal{A}$ be a positive relation algebra, then $\mathcal{R}$ is representable iff $(\mathcal{R}_{+})^{+}$ is completely representable.
\end{theorem}

\begin{theorem}
${\bf RPRA}$ is a canonical variety.
\end{theorem}

>>>>

\author{Daniel Rogozin}
\date{}
\title{Characterising representable positive relation algebras with Priestley duality}

@@

$ $

@@

\begin{fact} Let $\mathcal{L}$ be a distributive lattice and let $\mathcal{X}$ be a Priestley space:

\begin{enumerate}
\item $\mathcal{L} \hookrightarrow \mathcal{L}^{+} = (2^{\operatorname{Spec}(\mathcal{L})}, \cap, \cup, \emptyset, \operatorname{Spec}(\mathcal{L}))$,
\item $\mathcal{L} \cong \operatorname{ClOp}(\operatorname{Spec}(\mathcal{L}))$,
\item $\mathcal{X} \cong \operatorname{Spec}(\operatorname{ClOp}(\mathcal{X}))$,
\item The categories of Priestley spaces and bounded distributive lattices are dually equivalent.
\end{enumerate}
\end{fact}

\begin{enumerate}
\item $\mathcal{L}$ is completely representable iff $\mathcal{L}$ has a distinguishing set of complete, completely prime filters,
\item $(\mathcal{L}_+)^{+}$ is completely representable.
\end{enumerate}
\end{theorem}

@@

A positive relation algebra is \emph{completely representable} if it is representable and its bounded distributive lattice reduct is completely representable.

\section{Priestley spaces for positive relation algebras}

@@

\begin{fact} Let $\mathcal{R}$ be a positive relation algebra, then

\begin{enumerate}
\item If $F \subseteq \mathcal{R}$ is a filter, then for each $X, Y \subseteq \mathcal{R}$ $X ; Y \subseteq F$ iff $\uparrow (X ; Y) \subseteq F$
\item If $F_1, F_2 \subseteq \mathcal{R}$ are filters, so is $\uparrow (F_1 ; F_2)$,
\item Let $F_1, F_2$ be filters and $F_3 \in \operatorname{Spec}(\mathcal{R})$ such that $\uparrow (F_1 ; F_2) \subseteq F_3$, then there are $F_1', F_2' \in \operatorname{Spec}(\mathcal{R})$ such that $F_1 \subseteq F_1'$, $F_2 \subseteq F_2'$ and $\uparrow (F_1' ; F_2') \subseteq F_3$.
\end{enumerate}

\end{fact}

\begin{theorem} Let $\mathcal{R}$ be a positive relation algebra and $\mathcal{X}$ a PRA-space, then

\begin{enumerate}
\item $\mathcal{R} \hookrightarrow (2^{\operatorname{Spec}(R)}, \cup, \cap, \emptyset, \operatorname{Spec}(R), \bullet, \iota, \epsilon)$,
\item $\mathcal{R} \cong \operatorname{ClOp}(\operatorname{Spec}(\mathcal{R}))$,
\item $\mathcal{X} \cong \operatorname{Spec}(\operatorname{ClOp}(\mathcal{X}))$,
\item The categories of positive relation algebras and PRA-spaces are dually equivalent.
\end{enumerate}
\end{theorem}

@@

\begin{theorem}
Let $\mathcal{A}$ be a positive relation algebra, then $\mathcal{R}$ is representable iff $\mathcal{R}^{+}$ is completely representable.
\end{theorem}

\begin{proof}
The right-to-left implication is easy. If $\mathcal{R}^{+}$ is representable (no completeness needed here), so is $\operatorname{ClOp}(\operatorname{Spec}(\mathcal{L}))$ as a subalgebra of $\mathcal{R}^{+}$. But, by Priestley duality for positive relation algebras, $\operatorname{ClOp}(\operatorname{Spec}(\mathcal{L})) \cong \mathcal{L}$, so $\mathcal{L}$ is representable.

Assume that $\mathcal{R}$ is representable, then $T_{\mathcal{R}}$ is satisfiable, let $M \models T_{\mathcal{R}}$ and $M$ is $\omega$-saturated. Define a map $h : \mathcal{R}^+ \to 2^{M \times M}$ as
\begin{center}
$h : F \mapsto \{ (x, y) \in 1^M \: | \: f_{x, y} \in F \}$
\end{center}
where
\begin{center}
$f_{x, y} = \{ a \in \mathcal{R} \: | \: M \models R_a(x,y) \}$
\end{center}

\begin{claim}
$f_{x, y} \in \operatorname{Spec}(\mathcal{R})$ whenever $M \models 1(x, y)$.
\end{claim}

\begin{proof}
Let $a \in f_{x, y}$ and $a \leq b$.

Then $M \models R_a(x, y)$. But $a \leq b$ iff $a \cdot b = a$, so, from the axiom $R_{\cdot}(R_a, R_a, R_b)$ we have $M \models R_b(x, y)$.

Let $a, b \in f_{x, y}$, so $M \models R_a(x, y) \land R_b(x, y)$, so $R_{a \cdot b}(x, y)$ by the axiom $R_{\cdot}(R_a, R_a, R_b)$ again.

Let $a + b \in f_{x, y}$, so $M \models R_{a + b}(x, y)$, then we have either $M \models R_a(x, y)$ or $M \models R_b(x, y)$, so either $a \in f_{x,y}$ or $b \in f_{x,y}$.
\end{proof}

\begin{claim}
$h$ is one-to-one.
\end{claim}

\begin{claim}
$h$ is a complete representation of the bounded lattice reduct.
\end{claim}

\begin{proof}
Follows from the fact that $\operatorname{Spec}(\mathcal{L})$ is a Hausdorff space. (????)
\end{proof}

\begin{claim}
$h$ preserves the structure of an involutive monoid.
\end{claim}
\end{proof}

\begin{theorem}
${\bf RPRA}$ is a canonical variety. (is it a variety at all?)
\end{theorem}

\section{Union-free reducts of positive relation algebras}

TODO: \cite{bezhanishvili2011priestley}.
